---
title: "Significant Earthquakes, 1965-2016"
author: "Ying Liu"
date: "2021/7/17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here you will find some visualizations for the Earthquake Data Set. Have fun. :)


Used Libraries: 
```{r, warning=FALSE, message=FALSE}
library(lubridate)
library(xts)
library(ggplot2)
library(dplyr)
library(leaflet)
library(DT)
library(maps)
library(maptools)
library(viridisLite)
library(highcharter)
library(treemap)
library(viridisLite)
```


```{r, echo=FALSE}
setwd("/Users/LiuYing/Desktop/GU/2021-summer/assignment/assignment4/part2/")
quake<- read.csv("database.csv")
quake <- quake[,c("Date", "Time", "Latitude", "Longitude", "Type", "Depth", "Magnitude", "Magnitude.Type") ]
```

```{r, echo=FALSE}
quake$Date <- as.Date(quake$Date, format = "%m/%d/%Y")
quake$Year<- year(quake$Date)
quake$Month <- month(quake$Date)
quake$Day <- day(quake$Date)
head(quake)
```

##Mapping 

Mapping all earthquakes by using leaflet:
Nice tool for grouping the data points with the possibility to get more information about the earthquakes by zooming or clicking on one specific area.
```{r}
quake %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(lat=quake$Latitude, lng=quake$Longitude, clusterOptions = markerClusterOptions(),
             popup= paste(quake$Type,
        "<br><strong>Magnitude: </strong>", quake$Magnitude,
        "<br><strong>Depth: </strong>", quake$Depth,
        "<br><strong>Date: </strong>", quake$Date,
        "<br><strong>Date: </strong>", quake$Time
          ))
```


```{r}
quake %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lat=quake$Latitude, lng=quake$Longitude, weight=1, radius=1,
                   color= ifelse(quake$Magnitude>6.5,"red","yellow"),stroke=TRUE,
                   popup= paste(quake$Type,
        "<br><strong>Magnitude: </strong>", quake$Magnitude,
        "<br><strong>Depth: </strong>", quake$Depth,
        "<br><strong>Date: </strong>", quake$Date,
        "<br><strong>Date: </strong>", quake$Time)) %>%
  addLegend(labels=c("Magnitude > 6.5", "Magnitude < 6.5"), colors=c("red","yellow"))

```
In a next step, it would be nice to draw the faultines or plates above the map. 


##Affected Countries

Preparation: Converting Long Lat data to Country Names
```{r}
    world <- map('world', fill=TRUE, col="transparent", plot=FALSE)
    IDs <- sapply(strsplit(world$names, ":"), function(x) x[1])
    world_sp <- map2SpatialPolygons(world, IDs=IDs,
                     proj4string=CRS("+proj=longlat +datum=WGS84"))
    pointsSP <- SpatialPoints(cbind(x = quake$Longitude, y= quake$Latitude), 
                    proj4string=CRS("+proj=longlat +datum=WGS84"))
    indices <- over(pointsSP, world_sp)
    stateNames <- sapply(world_sp@polygons, function(x) x@ID)
    quake$Country <- stateNames[indices]
    
    quake_country <- quake[!is.na(quake$Country),]
```


Treemap showing the total amount of earthquakes occured per country: 
Indonesia had the most earthquakes over all years.
```{r, warning=FALSE}
sum_country <- quake_country %>%
 group_by(Country) %>%
summarise(Earthquakes=n())


sum_country %>%
  hchart("treemap", hcaes(x = Country, value = Earthquakes, color=Earthquakes)) %>%
  hc_credits(enabled = TRUE, style = list(fontSize = "10px")) %>%
  hc_title(text = "Earthquakes per Country")
  
```


Directory for Earthquakes per Country
```{r}
directory <- quake_country[, c("Country","Year","Month","Magnitude","Depth")] 
datatable(directory)
```



#Some more Plots
Earthquakes per Year and Month with ggplot

```{r}
quake<- quake[!is.na(quake$Date),]
```

```{r}
#Earthquakes per Year:
per_year <- quake %>%
  filter(Type=="Earthquake") %>%
  group_by(Year) %>%
summarise(Observations=n())


ggplot(per_year, aes(x=Year,y=Observations))+geom_bar(stat = "identity",fill="#58ACFA")+
labs(y="Observations",
      x="Year",
      title="Earthquakes per Year",
      caption="Source: Significant Earthquakes, 1965-2016")+
theme_grey()

#Earthquakes per Month:
per_month <- quake %>%
  filter(Type=="Earthquake") %>%
  group_by (Year, Month) %>%
  summarise(Observations=n())

per_month <- per_month %>%
  group_by (Month) %>%
  summarise(Mean=mean(Observations))

ggplot(per_month, aes(x=Month,y=Mean))+geom_bar(stat = "identity",fill="#58ACFA")+
labs(y="Average",
      x="Month",
      title="Average Earthquakes per Month",
      caption="Source: Significant Earthquakes, 1965-2016")+
theme_grey()


#Earthquakes per Day of a Month: 
per_day <- quake %>%
  filter(Type=="Earthquake") %>%
  group_by (Day,Year) %>%
  summarise(Observations=n())

per_day <- per_day %>%
  group_by (Day) %>%
  summarise(Mean=mean(Observations))

ggplot(per_day, aes(x=Day,y=Mean))+geom_bar(stat = "identity",fill="#58ACFA")+
labs(y="Mean",
      x="Day",
      title="Avergae Earthquakes per Day of a Month",
      caption="Source: Significant Earthquakes, 1965-2016")+
theme_grey()

```